{\large Gravitational wave sources in our Galactic backyard:}\\{\normalsize Predictions for BHBH, BHNS and NSNS binaries in \emph{LISA}

### 2021-11-12 paper/tex/sections/comparisons.tex
all Neige email comments done

\paragraph{\citet{Lau+2020}} focussed on the number of Galactic NSNS binaries that could be detected by LISA. Their study uses the same population synthesis code, COMPAS, as this work, though an earlier version. Despite this, their study finds a much larger number of detections. They make several different physical assumptions in their population synthesis, using the \citet{Fryer+2012} \textit{rapid} remnant mass prescription, limiting the maximum neutron star mass to $2 \unit{M_{\odot}}$ and not implementing PISN. However, we note that none of these assumptions strongly affect the NSNS LISA detection rate (see bottom panel of Fig.~\ref{fig:detection_rates}, models \modRapid{}, \modNSLow{} and \modNoPISN{}) and so this is unlikely to entirely account for our differences. It is also important to highlight that COMPAS has received several improvements and bug fixes since \citet{Vigna-Gomez+2018} (which contains the simulations used by \citet{Lau+2020}) and these could possibly have affected the formation rate of NSNSs.
%
Yet it is most likely that the remaining difference between our results is due to way in which we simulate the Milky Way. \citet{Lau+2020} use a model for the Milky Way similar to that of \citet{Breivik+2020}, which we use to estimate the impact of the choice of MW model is in Appendix~\ref{app:mw_changes}. We find that this simpler model for the Milky Way could result in an overestimation of the NSNS detection rate by at least a factor of two and so this may explain the discrepancy between our results. \citet{Lau+2020} additionally only uses a single metallicity rather than the two used in \citet{Breivik+2020} and so the overestimation could be even stronger as their results will have no contribution from lower metallicity systems.

\paragraph{\citet{Breivik+2020}} introduced the population synthesis code COSMIC and presented detections for many different DCO types in LISA using this code. They find that LISA will detect 93 BHBHs, 33 BHNSs and 8 NSNSs in the Milky Way over a 4 year mission. \citet{Breivik+2020} make many different physical assumptions from us, the most notable being that they assume the optimistic CE scenario and that case BB mass transfer is always unstable, whilst also using a simpler model for the Milky Way (see Appendix~\ref{app:mw_changes}). Thus for better comparison we ran our simulation using model \modCaseBBOpt{} and the Milky Way model from \citet{Breivik+2020}. This results in 97, 101 and 43 detections for BHBHs, BHNSs and NSNSs respectively. Therefore, though we are in very good agreement for BHBHs, we predict much higher rates for BHNSs and NSNSs. These differences are likely due to using a different population synthesis code (COSMIC), which has different underlying physics assumptions from COMPAS. Given our strong agreement for BHBHs, it is possible that COSMIC and COMPAS handle NSs differently and so lead to different detection rates for DCOs containing NSs. However checking this would require a more in-depth study of the intrinsic formation rate of DCOs containing NSs in the two codes.

@@

\noindent{}Overall, since the work of \citet{Nelemans+2001}, in additions to the LISA mission specifications, the methods that we use to simulate binaries and the Milky Way have all changed significantly. We now predict that LISA detections of these massive DCOs are dominated by BHBHs, rather than NSNSs, whilst the absolute detection rates for BHBHs and BHNSs are much higher. Further studies in this area could improve on this work by including the effects of systemic kicks on the position of systems in the Milky Way and accounting for contributions from other formation channels.

>>>>

\paragraph{\citet{Lau+2020}} focussed on the number of Galactic NSNS binaries that could be detected by LISA. Their study uses the same population synthesis code, COMPAS, as this work, though an earlier version. Despite this, their study finds a much larger number of detections. They make several different physical assumptions in their population synthesis, using the \citet{Fryer+2012} \textit{rapid} remnant mass prescription, limiting the maximum neutron star mass to $2 \unit{M_{\odot}}$ and not implementing PISN. However, we note that none of these assumptions strongly affect the NSNS LISA detection rate (see bottom panel of Fig.~\ref{fig:detection_rates}, models \modRapid{}, \modNSLow{} and \modNoPISN{}) and so this is unlikely to entirely account for the differences. It is also important to highlight that COMPAS has received several improvements and bug fixes since \citet{Vigna-Gomez+2018} (which contains the simulations used by \citet{Lau+2020}) and these could possibly have affected the formation rate of NSNSs.
%
Yet it is most likely that the remaining difference between our results is due to way in which we simulate the Milky Way. \citet{Lau+2020} use a model for the Milky Way similar to that of \citet{Breivik+2020}, which we use to estimate the impact of the choice of MW model in Appendix~\ref{app:mw_changes}. We find that this simpler model for the Milky Way could result in an overestimation of the NSNS detection rate by at least a factor of two and so this may explain the discrepancy between our results. \citet{Lau+2020} additionally only uses a single metallicity rather than the two used in \citet{Breivik+2020} and so the overestimation could be even stronger as their results will have no contribution from lower metallicity systems.

\paragraph{\citet{Breivik+2020}} introduced the population synthesis code COSMIC and presented detections for many different DCO types in LISA using this code. They find that LISA will detect 93 BHBHs, 33 BHNSs and 8 NSNSs in the Milky Way over a 4 year mission. \citet{Breivik+2020} make many physical assumptions that differ from our fiducial model, the most notable being that they assume the optimistic CE scenario and that case BB mass transfer is always unstable, whilst also using a simpler model for the Milky Way (see Appendix~\ref{app:mw_changes}). Thus for better comparison we ran our simulation using model \modCaseBBOpt{} and the Milky Way model from \citet{Breivik+2020}. This results in 97, 101 and 43 detections for BHBHs, BHNSs and NSNSs respectively. Therefore, though we are in very good agreement for BHBHs, we predict much higher rates for BHNSs and NSNSs. These differences are likely due to using a different population synthesis code (COSMIC), which has different underlying physics assumptions from COMPAS. Given our strong agreement for BHBHs, it is possible that COSMIC and COMPAS handle NSs differently and so lead to different detection rates for DCOs containing NSs. However checking this would require a more in-depth study of the intrinsic formation rate of DCOs containing NSs in the two codes.

@@

\noindent{}Overall, since the work of \citet{Nelemans+2001}, in addition to the LISA mission specifications, the methods that we use to simulate binaries and the Milky Way have all changed significantly. We now predict that LISA detections of these massive DCOs are dominated by BHBHs, rather than NSNSs, whilst the absolute detection rates for BHBHs and BHNSs are much higher. Further studies in this area could improve on this work by including the effects of systemic kicks on the position of systems in the Milky Way and accounting for contributions from other formation channels.

### 2021-11-12 paper/tex/sections/discussion.tex
all Neige email comments done

\subsubsection{Distinguishing from WDWD population}\label{sec:WDWD_distinguish}
The simplest way to check whether a source is a WDWD is to check its chirp mass. The mass of a non-rotating white dwarf cannot be larger than the Chandrasekhar limit of $1.4 \unit{M_\odot}$ \citep{Chandrasekhar+1931, Hamada+1961}, so we can take the maximum chirp mass of a WDWD to be $\sim 1.2 \unit{M_{\odot}}$. Therefore, any DCO with a chirp mass that satisfies $\mathcal{M}_c > 1.2 \unit{M_{\odot}} + \Delta \mathcal{M}_c$ must not be a WDWD (where $\Delta \mathcal{M}_c$ is calculated using Eq.~\ref{eq:chirp_mass_uncertainty}). We find that for the detectable population of a 4(10)-year LISA mission, \BHBHAboveMaxWDWDFourPerc{}(\BHBHAboveMaxWDWDTenPerc{})\% of BHBHs, \BHNSAboveMaxWDWDFourPerc{}(\BHNSAboveMaxWDWDTenPerc{})\% of BHNSs and \NSNSAboveMaxWDWDFourPerc{}(\NSNSAboveMaxWDWDTenPerc{})\% of NSNSs satisfy this condition. This method is not particularly effective for NSNSs since their average chirp mass, $1.17 \unit{M_\odot}$, is below the Chandrasekhar limit.

@@

\textit{Large progenitor masses:} Many of the black holes that are formed in our sample come from progenitors with large initial masses. The primary stars in systems that form  BHNSs and BHBHs have very primary progenitor masses that peak around 50\Msun and 80\Msun, but they extend well beyond that. The evolution of such massive progenitor stars is very poorly understood and observational constraints are very scarce in this mass range and practically non existent for the later rapid evolutionary phases.  This is reason for very substantial uncertainties in our final predictions which are unfortunately still hard to quantify at present. Additionally, the COMPAS code relies on fitting formula tat need to be extrapolated to create tracks are extrapolations at these higher masses.

@@

\textit{Limited metallicity range:} The stellar evolution fitting formulae that COMPAS uses is that they are limited to a metallicity range of $10^{-4} \le Z \le 0.03$ \citep{Hurley+2000} and should not be extrapolated too far outside this range. However, in our simulated Milky Way (based on the metallicity relation in \citealt{Frankel+2018}), the metallicity distribution can extend as far as $10^{-5} \le Z \le 0.06$, with a significant fraction of star formation occurring past $Z = 0.03$.

For metallicities below our range we assume they evolve similar to our lowest metallicity models. DCO formation at metallicities above $Z = 0.03$ is 
less efficient at high metallicity and so higher metallicities are unlikely to contribute significantly. Therefore, we place any sampled metallicity above the maximum of $Z = 0.03$ uniformly randomly in one of the top 5 highest bins that range across $0.01416 < Z < 0.03$ (since using a single metallicity for many binaries leads to unphysical artifacts).

>>>>

\subsubsection{Distinguishing from WDWD population}\label{sec:WDWD_distinguish}
The simplest way to check whether a source is a WDWD is to evaluate its chirp mass. The mass of a non-rotating white dwarf cannot be larger than the Chandrasekhar limit of $1.4 \unit{M_\odot}$ \citep{Chandrasekhar+1931, Hamada+1961}, so we can take the maximum chirp mass of a WDWD to be $\sim 1.2 \unit{M_{\odot}}$. Therefore, any DCO with a chirp mass that satisfies $\mathcal{M}_c > 1.2 \unit{M_{\odot}} + \Delta \mathcal{M}_c$ must not be a WDWD (where $\Delta \mathcal{M}_c$ is calculated using Eq.~\ref{eq:chirp_mass_uncertainty}). We find that for the detectable population of a 4(10)-year LISA mission, \BHBHAboveMaxWDWDFourPerc{}(\BHBHAboveMaxWDWDTenPerc{})\% of BHBHs, \BHNSAboveMaxWDWDFourPerc{}(\BHNSAboveMaxWDWDTenPerc{})\% of BHNSs and \NSNSAboveMaxWDWDFourPerc{}(\NSNSAboveMaxWDWDTenPerc{})\% of NSNSs satisfy this condition. This method is not particularly effective for NSNSs since their average chirp mass, $1.17 \unit{M_\odot}$, is below the Chandrasekhar limit.

@@

\textit{Large progenitor masses:} Many of the black holes that are formed in our sample come from progenitors with large initial masses. The primary stars in systems that form  BHNSs and BHBHs have very primary progenitor masses that peak around 50\Msun and 80\Msun, but they extend well beyond that. The evolution of such massive progenitor stars is very poorly understood and observational constraints are very scarce in this mass range and practically non existent for the later rapid evolutionary phases. This is reason for very substantial uncertainties in our final predictions which are unfortunately still hard to quantify at present. Additionally, the COMPAS code relies on fitting formula that are heavily extrapolated to be able to create evolutionary tracks for these higher mass progenitors.

@@

\textit{Limited metallicity range:} The stellar evolution fitting formulae that COMPAS uses is limited to a metallicity range of $10^{-4} \le Z \le 0.03$ \citep{Hurley+2000} and should not be extrapolated too far outside this range. However, in our simulated Milky Way (based on the metallicity relation in \citealt{Frankel+2018}), the metallicity distribution can extend as far as $10^{-5} \le Z \le 0.06$, with a significant fraction of star formation occurring past $Z = 0.03$.

For metallicities below the metallicity range allowed by COMPAS, we assume that binaries evolve similar to our lowest metallicity models. DCO formation at metallicities above $Z = 0.03$ becomes significantly less efficient for increasing metallicity. And so, higher metallicities are unlikely to contribute significantly. Therefore, we place any sampled metallicity above the maximum of $Z = 0.03$ uniformly randomly in one of the top 5 highest metallicity bins that range across $0.01416 < Z < 0.03$ (since using a single metallicity for many binaries leads to unphysical artifacts).
# (old text for easy comparison): 
#For metallicities below our range we assume they evolve similar to our lowest metallicity models. DCO formation at metallicities above $Z = 0.03$ is less efficient at high metallicity and so higher metallicities are unlikely to contribute significantly. Therefore, we place any sampled metallicity above the maximum of $Z = 0.03$ uniformly randomly in one of the top 5 highest bins that range across $0.01416 < Z < 0.03$ (since using a single metallicity for many binaries leads to unphysical artifacts).

### 2021-11-12 paper/tex/sections/introduction.tex
all Neige email comments done

Earlier work on BHBHs, BHNSs and NSNSs in LISA has used a variety of population synthesis codes, Milky Way models and LISA specifications, resulting in a wide range of predictions \citep{Nelemans+2001,Liu+2009,Belczynski+2010,Liu+2014,Lamberts+2019,Lau+2020,Breivik+2020,Sesana+2020}. We build upon previous efforts but with several important improvements. We explore the effects of varying binary physics assumptions by repeating our analysis for \nModels{} different models and comparing the effects on the detection rate and distributions of source parameters. We use a model for the Milky Way that accounts for the chemical enrichment history and is calibrated on the latest APOGEE survey \citep{Majewski+2017,Frankel+2018}, whereas most others did not consider the effect of metallicity in detail (see however \citealp{Lamberts+2019, Sesana+2020}). We provide a full treatment of the eccentricity of detectable sources both for the inspiral evolution as well as gravitational wave signal during the LISA mission. Moreover, our binary population synthesis simulation is the most extensive of its kind to date and makes use of the adaptive sampling algorithm STROOPWAFEL \citep{Broekgaarden+2019, Broekgaarden+2021}. Overall we simulate over 2 billion massive binaries to produce the DCO populations used in this work. We find that this large number of simulations is important to reduce the sampling noise even when using adaptive importance sampling.

All data produced in this study is publicly available on Zenodo \href{https://zenodo.org/record/4699713}{\faFileCode}\footnote{\url{https://zenodo.org/record/4699713}} as is the population used in our simulations \href{https://zenodo.org/record/4574727}{\faFileCode}\footnote{\url{https://zenodo.org/record/4574727}}. We make all code used to produce our results available in a Github repository \href{https://github.com/TomWagg/detecting-DCOs-in-LISA}{\faGithub}\footnote{\url{https://github.com/TomWagg/detecting-DCOs-in-LISA}}. In addition, the repository contains step-by-step Jupyter notebooks that explain how to reproduce and change each figure in the paper. In a companion paper, Wagg et al. 2021b (in preparation), we present \href{https://legwork.readthedocs.io}{\texttt{LEGWORK}}\footnote{\url{https://legwork.readthedocs.io}}, the \textbf{L}ISA \textbf{E}volution and \textbf{G}ravitational \textbf{W}ave \textbf{Or}bit \textbf{K}it, a python package designed for making predictions for the detection of sources with LISA, which we use in this work.

>>>>

Earlier work on BHBHs, BHNSs and NSNSs in LISA has used a variety of population synthesis codes, Milky Way models and LISA specifications, resulting in a wide range of predictions \citep{Nelemans+2001,Liu+2009,Belczynski+2010,Liu+2014,Lamberts+2019,Lau+2020,Breivik+2020,Sesana+2020}. We build upon previous efforts but with several important improvements. We explore the effects of varying binary physics assumptions by repeating our analysis for \nModels{} different models and comparing the effects on the detection rate and distributions of source parameters. We use a model for the Milky Way that accounts for the chemical enrichment history and is calibrated on the APOGEE survey \citep{Majewski+2017,Frankel+2018}, whereas most others did not consider the effect of metallicity in detail (see however \citealp{Lamberts+2019, Sesana+2020}). We provide a full treatment of the eccentricity of detectable sources both for the inspiral evolution as well as gravitational wave signal during the LISA mission. Moreover, our binary population synthesis simulation is the most extensive of its kind to date and makes use of the adaptive sampling algorithm STROOPWAFEL \citep{Broekgaarden+2019, Broekgaarden+2021}. Overall we simulate over 2 billion massive binaries to produce the DCO populations used in this work. We find that this large number of simulations is important to reduce the sampling noise even when using adaptive importance sampling.

All data related to the predictions made in this study are publicly available on Zenodo at \citet{Wagg+2021_zenodo}, as are the populations used in our simulations at \citet[][BHBH]{Broekgaarden:2021-zenodo-BHBH} \citet[][BHNS]{Broekgaarden:2021-zenodo-BHNS} and  \citet[][NSNS]{Broekgaarden:2021-zenodo-NSNS}. We make all code used to produce our results available in a Github repository \href{https://github.com/TomWagg/detecting-DCOs-in-LISA}{\faGithub}\footnote{\url{https://github.com/TomWagg/detecting-DCOs-in-LISA}}. In addition, the repository contains step-by-step Jupyter notebooks that explain how to reproduce and change each figure in the paper. In a companion paper, Wagg et al. 2021b (in preparation), we present \href{https://legwork.readthedocs.io}{\texttt{LEGWORK}}\footnote{\url{https://legwork.readthedocs.io}}, the \textbf{L}ISA \textbf{E}volution and \textbf{G}ravitational \textbf{W}ave \textbf{Or}bit \textbf{K}it, a python package designed for making predictions for the detection of sources with LISA, which we use in this work.

### 2021-11-12 paper/tex/sections/methods.tex
all Neige email comments done

The result of the simulations is a sample of binaries, which, for each metallicity $Z$, have $N_{\rm binaries}$ binaries with parameters
\begin{equation}
    \mathbf{b}_{{Z, i}} = \{m_1, m_2, a_{\rm DCO}, e_{\rm DCO}, t_{\rm evolve}, t_{\rm inspiral}, w\},
\end{equation}
for $i = 1, 2, \dots, N_{\rm binaries}$, where $m_1$ and $m_2$ are the primary and secondary masses, $a_{\rm DCO}$ and $e_{\rm DCO}$ are the semi-major axis and eccentricity at the moment of double compact object (DCO) formation, $t_{\rm evolve}$ is the time between the binary's zero-age main sequence and DCO formation, $t_{\rm inspiral}$ is the time between DCO formation (that is immediately after the second supernova in the system) and gravitational wave merger and $w$ is the adaptive importance sampling weight assigned by STROOPWAFEL \cite[][Eq.~7]{Broekgaarden+2019}. We sample from these sets of parameters when creating synthetic galaxies.

@@

Most previous studies that predict a detection rate for LISA place binaries in the Milky Way independently of their age or evolution. We improve upon this as the first study to use an empirically-informed analytical model of the Milky Way that takes into account the galaxy's enrichment history by applying the metallicity-radius-time relation from \citet{Frankel+2018}. The authors developed this relation in order to measure the global efficiency of radial migration in the Milky Way and calibrated it using a sample of red clump stars measured with APOGEE \citep{Majewski+2017}.

@@

Our model for the Milky Way accounts for the low-\achem\footnote{Nomenclature used to describe the enhancement of $\alpha$ elements compared to iron in stellar atmospheres}~disc, high-\achem~disc and a central component approximating a bar/bulge. The low- and high-\achem~discs are often also referred to as the thin and thick discs because the stellar vertical distribution is better fit by a double exponential rather than a single one. However, this doesn't allow one to assign a star to either the thin or thick disk purely based on its height above the Galactic plane. Therefore, we instead use the chemical definition of the two disks (applying the \achem~nomenclature) as there is a clear bimodal distribution in the chemical plane, allowing stars to be assigned to each of the disc components based on their chemical abundances. For each of the three components, we use a separate star formation history and spatial distribution, which we combine into a single model, weighting each component by its stellar mass. \citet{Licquia+2015} gives that the stellar mass of the bulge is $0.9 \times 10^{10} \unit{M_{\odot}}$ and the stellar mass of the disc is $5.2 \times 10^{10} \unit{M_\odot}$, which we split equally between the low- and high-\achem~discs \citep[e.g.,][]{Snaith+2014}.

@@

\textit{Radial distribution:} For each of the three components we employ the same single exponential distribution (but with different scale lengths)
\begin{equation}\label{eq:galaxy_R}
    p(R) = \exp(-\frac{R}{R_d}) \frac{R}{R_d^2},
\end{equation}
where $R$ is the Galactocentric radius and $R_d$ is the scale length of the component. For the low-\achem~disc, we set $R_d = R_{\rm exp}(\tau)$, where $R_{\rm exp}(\tau)$ is the scale length presented in \citet[][Eq.~5]{Frankel+2018}
\begin{equation}
    R_{\rm exp}(\tau) = 4 \unit{kpc} \qty(1 - \alpha_{R_{\rm exp}} \qty(\frac{\tau}{8 \unit{Gyr}})),
\end{equation}
where $\alpha_{R_{\rm exp}} = 0.3$ is the inside-out growth parameter\footnote{In $R_{\rm exp}(\tau)$, we use 4 kpc instead of 3 kpc for the 0 Gyr exponential scale-length of the disc as NF finds that it provides a better fit to the original data}. This scale length accounts for the inside-out growth of the low-\achem~disc and hence is age dependent. We assume $R_d = (1 / 0.43) \unit{kpc}$ for the high-\achem~disc \citep[][Table~1]{Bovy+2016} and $R_d = 1.5 \unit{kpc}$ for the bar component \citep{Bovy+2019}. We show the combination of these distributions in the second panel on the left in Fig.~\ref{fig:galaxy_schematic}.

@@

\textit{Metallicity-radius-time relation:} To account for the chemical enrichment of star forming gas as the Milky Way evolves, we adopt the relation given by \citep[][Eq. 7]{Frankel+2018}
\begin{equation}\label{eq:galaxy_FeH}
    \begin{split}
        [{\rm Fe} / {\rm H}] (R, \tau) &= F_m + \nabla [{\rm Fe} / {\rm H}] R \\
        &- \qty(F_m + \nabla [{\rm Fe} / {\rm H}] R^{\rm now}_{[{\rm Fe} / {\rm H}] = 0} ) f(\tau),
    \end{split}
\end{equation}
where
\begin{equation}
    f(\tau) = \qty(1 - \frac{\tau}{\tau_m})^{\gamma_{[{\rm Fe} / {\rm H}]}},
\end{equation}
$F_m = -1 \unit{dex}$ is the metallicity of the gas at the center of the disc at $\tau = \tau_m$, $\nabla [{\rm Fe} / {\rm H}] = -0.075 \unit{kpc^{-1}}$ is the metallicity gradient, $R^{\rm now}_{[{\rm Fe} / {\rm H}] = 0} = 8.7 \unit{kpc}$ is the radius at which the present day metallicity is solar and $\gamma_{[{\rm Fe} / {\rm H}]} = 0.3$ set the time dependence of the chemical enrichment. We can convert this to the representation of metallicity that we use in this paper by applying \citep[e.g][]{Bertelli+1994}
\begin{equation}\label{eq:galaxy_FeH_to_Z}
    \log_{10} (Z) = 0.977 [{\rm Fe} / {\rm H}] + \log_{10}(Z_\odot).
\end{equation}

@@

For each Milky Way instance, we randomly sample the following set of parameters
\begin{equation}
    \mathbf{g}_{{j}} = \{\tau, R, Z, z, \theta\}
\end{equation}
for $j = 1, 2, \dots, N_{\rm MW}$, where we set $N_{\rm MW} = 2 \times 10^{5}$, $\tau, R, Z$ and $z$ are defined and sampled using the distribution functions specified in Section~\ref{sec:mw_model}, $\theta$ is the polar angle sampled uniformly on $[0, 2\pi)$ and $Z$ is the metallicity. Figure~\ref{fig:galaxy_schematic} shows an example of a random Milky Way instance created with these distributions. This shows how these distributions translate to positions and illustrates the gradient in metallicity over radius.

@@

The magnitude of the systemic kicks are typically small compared to the initial circular velocity of a binary at each Galactocentric radius. Therefore, we expect that kicks will not significantly alter the overall distribution of their positions (see however e.g.\ \citealt{Brandt+1995, Abbott+2017_GW170817_progenitor}). Given this, and for the sake of computational efficiency, we do not account for the displacement due to systemic kicks in our analysis.

@@

Note that the length of the LISA mission affects how much a binary will evolve over the mission and thus changes the frequency limits in Eq.~\ref{eq:snr}. The SNR generally scales as $\sqrt{T_{\rm obs}}$ (with exceptions for sources very close to merging) and thus the SNR of a typical source in a 10-year LISA mission is approximately $1.58$ (=$\sqrt{10/4}$) times stronger than in a 4-year mission.

We use \href{https://legwork.readthedocs.io/en/latest/}{LEGWORK} to calculate the signal-to-noise ratio for each binary and the package ensures that enough harmonics are computed for each binary such that the error on the gravitational wave luminosity remains below 1\%.

>>>>

The result of the simulations is a sample of binaries, which, for each metallicity $Z$, have $N_{\rm binaries}$ binaries with parameters
\begin{equation}
    \mathbf{b}_{{Z, i}} = \{m_1, m_2, a_{\rm DCO}, e_{\rm DCO}, t_{\rm evolve}, t_{\rm inspiral}, w\},
\end{equation}
for $i = 1, 2, \dots, N_{\rm binaries}$, where $m_1$ and $m_2$ are the primary and secondary masses, $a_{\rm DCO}$ and $e_{\rm DCO}$ are the semi-major axis and eccentricity at the moment of double compact object (DCO) formation, $t_{\rm evolve}$ is the time between the binary's zero-age main sequence and DCO formation, $t_{\rm inspiral}$ is the time between DCO formation (that is immediately after the second supernova in the system) and gravitational-wave merger and $w$ is the adaptive importance sampling weight assigned by STROOPWAFEL \cite[][Eq.~7]{Broekgaarden+2019}. We sample from these sets of parameters when creating synthetic galaxies.

@@

Most previous studies that predict a detection rate for LISA place binaries in the Milky Way independently of their age or evolution. We improve upon this as the first study to use an empirically-informed analytical model of the Milky Way that takes into account the galaxy's enrichment history by applying the metallicity-radius-time relation from \citet{Frankel+2018}. Those authors developed this relation in order to measure the global efficiency of radial migration in the Milky Way and calibrated it using a sample of red clump stars measured with APOGEE \citep{Majewski+2017}.

@@

Our model for the Milky Way accounts for the low-\achem\footnote{Nomenclature used to describe the enhancement of $\alpha$ elements compared to iron in stellar atmospheres}~disc, high-\achem~disc and a central component approximating a bar/bulge. The low- and high-\achem~discs are often also referred to as the thin and thick discs because the stellar vertical distribution is better fit by a double exponential rather than a single one. However, this doesn't allow one to assign a star to either the thin or thick disk purely based on its height above the Galactic plane. Therefore, we instead use the chemical definition of the two disks (applying the \achem~nomenclature) as there is a clear bimodal distribution in the chemical plane, allowing stars to be assigned to each of the disc components based on their chemical abundances. For each of the three components, we use a separate star formation history and spatial distribution, which we combine into a single model, weighting each component by its present-day stellar mass. \citet{Licquia+2015} gives that the stellar mass of the bulge is $0.9 \times 10^{10} \unit{M_{\odot}}$ and the stellar mass of the disc is $5.2 \times 10^{10} \unit{M_\odot}$, which we split equally between the low- and high-\achem~discs \citep[e.g.,][]{Snaith+2014}.

@@

\textit{Radial distribution:} For each of the three components we employ the same single exponential distribution (but with different scale lengths)
\begin{equation}\label{eq:galaxy_R}
    p(R) = \exp(-\frac{R}{R_d}) \frac{R}{R_d^2},
\end{equation}
where $R$ is the Galactocentric radius and $R_d$ is the scale length of the component. For the low-\achem~disc, we set $R_d = R_{\rm exp}(\tau)$, where $R_{\rm exp}(\tau)$ is the scale length presented in \citet[][Eq.~6]{Frankel+2018}
\begin{equation}
    R_{\rm exp}(\tau) = 4 \unit{kpc} \qty(1 - \alpha_{R_{\rm exp}} \qty(\frac{\tau}{8 \unit{Gyr}})),
\end{equation}
where $\alpha_{R_{\rm exp}} = 0.3$ is the inside-out growth parameter\footnote{We find that $R_{\rm exp}(\tau) = 4$ kpc fits the data well and adopt this value rather than the 3 kpc quoted in \cite{Frankel+2018}, which was a fixed parameter (not a fit).}

This scale length accounts for the inside-out growth of the low-\achem~disc and hence is age dependent. We assume $R_d = (1 / 0.43) \unit{kpc}$ for the high-\achem~disc \citep[][Table~1]{Bovy+2016} and $R_d = 1.5 \unit{kpc}$ for the bar component \citep{Bovy+2019}. We show the combination of these distributions in the second panel on the left in Fig.~\ref{fig:galaxy_schematic}.

@@

\textit{Metallicity-radius-time relation:} To account for the chemical enrichment of star forming gas as the Milky Way evolves, we adopt the relation given by \citep[][Eq. 7]{Frankel+2018}
\begin{equation}\label{eq:galaxy_FeH}
    \begin{split}
        [{\rm Fe} / {\rm H}] (R, \tau) &= F_m + \nabla [{\rm Fe} / {\rm H}] R \\
        &- \qty(F_m + \nabla [{\rm Fe} / {\rm H}] R^{\rm now}_{[{\rm Fe} / {\rm H}] = 0} ) f(\tau),
    \end{split}
\end{equation}
where
\begin{equation}
    f(\tau) = \qty(1 - \frac{\tau}{\tau_m})^{\gamma_{[{\rm Fe} / {\rm H}]}},
\end{equation}
$F_m = -1 \unit{dex}$ is the metallicity of the gas at the center of the disc at $\tau = \tau_m$, $\nabla [{\rm Fe} / {\rm H}] = -0.075 \unit{kpc^{-1}}$ is the metallicity gradient, $R^{\rm now}_{[{\rm Fe} / {\rm H}] = 0} = 8.7 \unit{kpc}$ is the radius at which the present day metallicity is solar and $\gamma_{[{\rm Fe} / {\rm H}]} = 0.3$ sets the time dependence of the chemical enrichment. We can convert this to the representation of metallicity that we use in this paper by applying \citep[e.g.][]{Bertelli+1994}
\begin{equation}\label{eq:galaxy_FeH_to_Z}
    \log_{10} (Z) = 0.977 [{\rm Fe} / {\rm H}] + \log_{10}(Z_\odot).
\end{equation}

@@

For each Milky Way instance, we randomly sample the following set of parameters
\begin{equation}
    \mathbf{g}_{{j}} = \{\tau, R, Z, z, \theta\}
\end{equation}
for $j = 1, 2, \dots, N_{\rm MW}$, where we set $N_{\rm MW} = 2 \times 10^{5}$, $\tau, R, Z$ and $z$ are defined and sampled using the distribution functions specified in Section~\ref{sec:mw_model}, $\theta$ is the azimuthal angle sampled uniformly on $[0, 2\pi)$ and $Z$ is the metallicity. Figure~\ref{fig:galaxy_schematic} shows an example of a random Milky Way instance created with these distributions. This shows how these distributions translate to positions and illustrates the gradient in metallicity over radius.

@@

The magnitude of the systemic kicks are typically small compared to the initial circular velocity of a binary at each Galactocentric radius. Therefore, we expect that kicks will not significantly alter the overall distribution of their positions (see however, e.g., \citealt{Brandt+1995, Abbott+2017_GW170817_progenitor}). Given this, and for the sake of computational efficiency, we do not account for the displacement due to systemic kicks in our analysis.

@@

Note that increasing the length of the LISA mission allows more time for a DCO to evolve over the mission. Therefore the frequency limits in Eq.~\ref{eq:snr} are dictated by the LISA mission length. The SNR generally scales as $\sqrt{T_{\rm obs}}$ (with exceptions for sources very close to merging) and thus the SNR of a typical source in a 10-year LISA mission is approximately $1.58$ (=$\sqrt{10/4}$) times stronger than in a 4-year mission.

We use \href{https://legwork.readthedocs.io/en/latest/}{LEGWORK} to calculate the signal-to-noise ratio for each binary and the package ensures that enough harmonics are computed for each binary such that the error on the gravitational-wave luminosity remains below 1\%.
